R Notebook

### 2023-11-25 R_code/paper_analysis.Rmd
subsampling validation adding brier score fixing stuff

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
```
```{r}
validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
```

@@

Choose specific program
```{r}
program = 'iqtree'
if (program=='fasttree') {
    all_models_performance = fasttree_all_models_performance
    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
    group_performance = fasttree_group_performance
    vi = fasttree_vi
    enriched_test = fasttree_enriched_test
    #validation_data = fasttree_validation_data

@@

summm<- enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot)

                                                )

summm%>% ggplot(aes(x=mean_true,y=mean_bayes))+geom_point()+geom_abline(slope=1, intercept = 0)
summm%>% ggplot(aes(x=mean_true,y=mean_alrt))+geom_point()+geom_abline(slope=1, intercept = 0)

summm %>% ggplot(aes(x=mean_true,y=mean_bayes_cal))+geom_point()+geom_abline(slope=1, intercept = 0)
summm %>% ggplot(aes(x=mean_true,y=mean_alrt_cal))+geom_point()+geom_abline(slope=1, intercept = 0)

summm  %>% ggplot(aes(x=mean_true,y=mean_pred))+geom_point()+geom_abline(slope=1, intercept = 0)

pivoted<- summ %>% pivot_longer()
```

General performance
```{r}

out_path <-paste(prefix,"ML_results/",program,"/performance_text",".tsv", sep = "")

paper_performance<-final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance
write_tsv(paper_performance,out_path)

@@

Variable importance 
```{r}
out_path <-paste(prefix,"ML_results/",program,"/vi_text",".tsv", sep = "")

vi_data<- vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

@@

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',

@@

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc ) %>% arrange(feature_name,-`Gini-importance`)
vi_data
write_tsv(vi_data,out_path)

@@

@@

```{r}
ggplotConfusionMatrix <- function(m){
  mytitle <- ""#paste("Accuracy", percent_format()(m$overall[1]))
                  # "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = scales::comma(Freq))) +
    theme(legend.position = "none") +
    ggtitle(mytitle)+theme(text = element_text(size = 13))
  return(p)
}
conf_data<-enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)

@@

```
Prediction metrics

```{r}
library(predtools)
library(MASS)
library(caret)

@@

```
```{r}
test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') 
```

@@

@@

```{r}
cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)
```

@@

running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
)) 
summary(running_time_data %>% pull (diff1))
summary(running_time_data %>% pull (diff2))
summary(running_time_data %>% pull (extraction_running_time))
summary(running_time_data %>% pull (nni_ll_eval_running_time))
summary(running_time_data %>% pull (program_boot_run_time))
#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
```
```{r}
summary(validation_data$feature_n_unique_seq)
raxml_validation_data$
```

>>>>

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
library(predtools)
library(MASS)
library(caret)
```
```{r}
validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
```

@@

Choose specific program
```{r}
program = 'fasttree'
if (program=='fasttree') {
    all_models_performance = fasttree_all_models_performance
    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
    group_performance = fasttree_group_performance
    vi = fasttree_vi
    enriched_test = fasttree_enriched_test
    #validation_data = fasttree_validation_data

@@

IQTREE
```{r}

summm_iqtree<- iqtree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_ultrafast = mean(bootstrap_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot))

bayes<-summm_iqtree%>% ggplot(aes(y=mean_true,x=mean_bayes))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aBayes test")+theme(text = element_text(size = 11))+ coord_flip()

alrt<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_alrt))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aLRT test ")+theme(text = element_text(size = 11))+ coord_flip()

ultrafast<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_ultrafast/100))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Ultrafast\nbootstrap")+theme(text = element_text(size = 11))+ coord_flip()

pred_iqtree<-summm_iqtree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("IQTREE\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

iqtree_plot<-ggarrange(pred_iqtree,ultrafast,bayes,alrt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )
iqtree_plot
```
FASTTREE

```{r}
summm_fasttree<- fasttree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))

pred_fasttree<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Fasttree\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

sh<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("SH test")+theme(text = element_text(size = 11))+ coord_flip()

fasttree_plot<-ggarrange(pred_fasttree,sh, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
fasttree_plot
```

RAXML
```{r}
summm_raxml<- raxml_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))

pred_raxml<-summm_raxml  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0, color = "green")+ylab("True fraction")+xlab("RAxML-NG\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

rapid<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Bootstrap\nsupport")+theme(text = element_text(size = 11))+ coord_flip()

raxml_plot<-ggarrange(pred_raxml,rapid, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
raxml_plot
```

```{r}
ggarrange(pred_iqtree,alrt, bayes, ultrafast,pred_raxml,rapid,"","",pred_fasttree,sh,align = "h", nrow = 3, ncol = 4 )

@@

General performance
IQTREE
```{r}

out_path <-paste(prefix,"ML_results/","iqtree","/performance_text",".tsv", sep = "")

paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance_iqtree
write_tsv(paper_performance_iqtree,out_path)

paper_performance_iqtree %>% arrange(description)

```
FASTTREE
```{r}

out_path <-paste(prefix,"ML_results/","fasttree","/performance_text",".tsv", sep = "")

paper_performance_fasttree<-fasttree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw SH support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated SH support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ SH support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance_fasttree
write_tsv(paper_performance_fasttree,out_path)
paper_performance_fasttree %>% arrange(description)

```
RAXML
```{r}

out_path <-paste(prefix,"ML_results/","raxml","/performance_text",".tsv", sep = "")

paper_performance_raxml<-raxml_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance_raxml
write_tsv(paper_performance_raxml,out_path)
paper_performance_raxml %>% arrange(description)

Results on test data
```{r}
total_model_performance = rbind(paper_performance_iqtree %>% mutate(program='IQTREE'), paper_performance_fasttree %>% mutate(program='FastTree'), paper_performance_raxml %>% mutate(program='RAxML-NG')) %>% filter (dataset %in% c("test"), description %in% c("all features","Raw SH support","Raw aLRT support","Raw aBayes support","Raw bootstrap support","Raw ultrafast bootstrap support")) %>% mutate(Support = description) %>% rename(MCC = mcc_0.5)

mcc_plot<-total_model_performance %>% ggplot(aes(y=MCC, x = program, fill = Support))+geom_col(position = "dodge")
logloss_plot<-total_model_performance %>% ggplot(aes(y=logloss, x = program, fill = Support))+geom_col(position = "dodge")
#total_model_performance %>% ggplot(aes)

ggarrange(mcc_plot,logloss_plot, labels = c("A","B"),align = "h", nrow = 2, ncol = 1, common.legend = TRUE )
```

Variable importance 
IQTREE

```{r}
out_path <-paste(prefix,"ML_results/","iqtree","/vi_text",".tsv", sep = "")

vi_data<- iqtree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

@@

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

@@

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`,feature_name)
vi_data
write_tsv(vi_data,out_path)

@@

FASTTREE

```{r}
out_path <-paste(prefix,"ML_results/","fasttree","/vi_text",".tsv", sep = "")

vi_data<- fasttree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data 
write_tsv(vi_data,out_path)
```

RAXML

```{r}
out_path <-paste(prefix,"ML_results/","raxml","/vi_text",".tsv", sep = "")

vi_data<- raxml_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data %>% arrange(feature_name)
write_tsv(vi_data,out_path)
```

@@

```{r}
ggplotConfusionMatrix <- function(m){
  mytitle <- ""#paste("Accuracy", percent_format()(m$overall[1]))
                  # "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = scales::comma(Freq))) +
    theme(legend.position = "none") +
    ggtitle(mytitle)+theme(text = element_text(size = 13))
  return(p)
}
conf_data<-iqtree_enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)

@@

@@

@@

cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)

@@

Running time

@@

running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
)) 
summary(running_time_data %>% pull (diff1))
summary(running_time_data %>% pull (diff2))
summary(running_time_data %>% pull (extraction_running_time))
summary(running_time_data %>% pull (nni_ll_eval_running_time))
summary(running_time_data %>% pull (program_boot_run_time))
#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
```

```{r}
summary(validation_data$feature_n_unique_seq)
raxml_validation_data$
```
